smooth-sea

### 2018-05-22 index.rmd
first draft complete

full_panel <- tract %>%
  group_by(GISJOIN) %>% 
  summarize(n = n()) %>%
  pull(n) %>%
  max

@@

#periods <- tract %>%
#  filter(listingQtr != as.yearqtr(Sys.Date())) %>%
#  summarize(n = length(unique(listingQtr))) %>% 
#  pull(n)

@@

#for each unique quarter in the CL data
for(i in unique(as.character(tract$listingQtr))){

  period <- tract %>% #start pipe with tract, end with period
    mutate(listingQtr = factor(listingQtr, ordered = T)) %>% #make factor
    filter(listingQtr == i) %>% #filter to qtr i
    right_join(census) %>% #right join to census, allows missingness
    ungroup() %>% #remove grouping by tract
    mutate(listingQtr = i, #listingQtr == current iteration
           Qtr = gsub(pattern = "^\\d{4} ", replacement = "", x = listingQtr)) #subset str to Q1,2,3,4
  panel <- bind_rows(panel, period) #append to panel object
}

@@

summary(m.nsar1) #NB: does not improve fit to use random effect time specification, using fixed effects
sea_df$nsar1_Med <- m.nsar1$summary.linear.predictor[, "0.5quant"]
sea_df$nsar1_SD <- m.nsar1$summary.linear.predictor[, "sd"]
sea_df$nsar1_postWidth <- m.nsar1$summary.linear.predictor[, "0.975quant"] - m.nsar1$summary.linear.predictor[,"0.025quant"]
sea_df$nsar1_Eff <- m.nsar1$summary.random$idtract[sea_df$idtract, "0.5quant"]

@@

#spatio-temporal smoothing model with linear space-time interaction (Bernadelli 1995)
form.spt <- log(med1B) ~ 1 + Qtr +
  f(idtract, model = "bym", #ICAR spatial RE for tract neighbors + IID RE for tract
    scale.model = T,
    graph = "R:/Project/seattle_rental_market/report/spatial_epi/seatract.graph") +
  f(idtractqtr, model="iid") + #exchangeable RE for period
  f(idqtr, model = "rw1")

@@

This website provides information for the paper _Estimating Neighborhood Asking Rents using Scraped Craigslist Rental Listings_

You can download the R source code and data for this project from Github [here](https://github.com/hesscl/smooth-sea)

@@

@@

```{r index.maps, echo=F}
obs <- sea_f %>%
  mutate(rent = actualRent) %>%
  filter(listingQtr == as.yearqtr(Sys.Date())) %>%
  mutate(var = paste(as.yearqtr(Sys.Date()), "to-Date"))

pred <- sea_f %>%
  mutate(rent = nsar1_Med) %>%
  filter(listingQtr == as.yearqtr(Sys.Date())) %>%
  mutate(var = paste(as.yearqtr(Sys.Date()), "Forecast"))

comb <- bind_rows(obs, pred) %>%
  mutate(tooltip = paste0(NAMELSAD10, "<br>", "Forecast: $", round(nsar1_Med), "<br>", "Qtr-to-Date: $", round(actualRent), "<br>", "N Qtr-to-Date: ", n1B))
comb <- SharedData$new(comb, ~GISJOIN)

@@

ggplotly(avp_map, tooltip = c("label")) %>%
  layout(autosize = T) %>%
  highlight(on = "plotly_hover", off = NULL)
```

@@

```{r, echo=F}
gg_hist <- ggplot(sea_df, aes(x = med1B, group = listingQtr, color = listingQtr)) +
  geom_density() +
  theme_minimal() +
  scale_color_viridis_d(end = .8)

ggplotly(gg_hist) %>%
  layout(autosize = T)
```

@@

int_df <- sea_df %>%
  group_by(listingQtr) %>%
  mutate(value = int_Med,
         qtrDev = value-mean(value, na.rm=T),
         model = "B. Quarter FE")

ns_df <- sea_df %>%
  group_by(listingQtr) %>%
  mutate(value = ns_Med,
         qtrDev = value-mean(value, na.rm=T),
         model = "C. Non-Spatial RE")

nsar1_df <- sea_df %>%
  group_by(listingQtr) %>%
  mutate(value = nsar1_Med,
         qtrDev = value-mean(value, na.rm=T),
         model = "D. Non-Spatial AR(1) RE")

bym_df <- sea_df %>%
  group_by(listingQtr) %>%
  mutate(value = bym_Med,
         qtrDev = value-mean(value, na.rm=T),
         model = "E. Spatial RE")

spt_df <- sea_df %>%
  group_by(listingQtr) %>%
  mutate(value =  spt_Med,
         qtrDev = value-mean(value, na.rm=T),
         model = "F. Spatiotemporal")

@@

obs_lineplot <- ggplot(panel_df, aes(x = listingQtr, y = value, group = GISJOIN, label = GISJOIN, color = qtrDev, frame = model)) +
  geom_line(lwd = .25, color = "black", alpha = .5) +
  geom_point() +
  scale_color_gradient2_tableau()

ggplotly(obs_lineplot, tooltip = c("label", "y")) %>%
  layout(autosize = T) %>%
  animation_opts(frame = 1000, transition = 0, redraw = FALSE) %>%
  animation_button(hide = T) %>%
  highlight(on = "plotly_hover", off = NULL)
```

@@

```{r spat.series, echo=F}
map_panel <- SharedData$new(sea_f, ~GISJOIN)

map_panel_gg <- ggplot(map_panel, aes(x = long, y = lat, group = group, fill = actualRent, label = GISJOIN, frame = listingQtr)) +
  geom_polygon(data=sea_f, aes(fill = NULL, label = NULL), fill = "grey80") +
  geom_polygon(color = "grey85", lwd = .05) +
  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey80") +
  coord_quickmap() +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_blank()) +
  labs(fill = "Median<br>1B Rent")

ggplotly(map_panel_gg, tooltip = c("label", "fill")) %>%
  layout(autosize = T) %>%
  animation_opts(frame = 1000, transition = 0, redraw = FALSE)
```

@@

Column {data-width=400}
-------------------------------------

@@

Column {data-width=600}
-------------------------------------

### Actual vs. Predicted Scatterplot

```{r, echo=F}
avp_scatter <- ggplot(sea_df, aes(x = actualRent, y = bym_Med, size = n1B, frame = listingQtr)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1) +
  scale_size_area() +
  coord_equal() +
  theme_minimal()

ggplotly(avp_scatter) %>%
  layout(autosize = T)
```

Forecast
=====================================

>>>>

@@

@@

#for each unique quarter in the CL data
for(i in unique(as.character(tract$listingQtr))){
  period <- tract %>% #start pipe with tract, end with period
    mutate(listingQtr = factor(listingQtr, ordered = T)) %>% #make factor
    filter(listingQtr == i) %>% #filter to qtr i
    right_join(census) %>% #right join to census, allows missingness
    ungroup() %>% #remove grouping by tract
    mutate(listingQtr = i, #listingQtr == current iteration
           Qtr = gsub(pattern = "^\\d{4} ", replacement = "", x = listingQtr)) #subset str to Q1,2,3,4
  panel <- bind_rows(panel, period) #append to panel object
}

@@

summary(m.nsar1) #NB: does not improve fit to use random effect time specification, using fixed effects
sea_df$nsar1_Med <- m.nsar1$summary.linear.predictor[, "0.5quant"]
sea_df$nsar1_SD <- m.nsar1$summary.linear.predictor[, "sd"]
sea_df$nsar1_postWidth <- m.nsar1$summary.linear.predictor[, "0.975quant"] - m.nsar1$summary.linear.predictor[,"0.025quant"]
sea_df$nsar1_Eff <- m.nsar1$summary.random$idqtr[sea_df$idqtr, "0.5quant"]

@@

#spatio-temporal smoothing model, AR(1) process with linear space-time interaction (Bernadelli 1995)
form.spt <- log(med1B) ~ 1 +
  f(idtract, model = "bym", #ICAR spatial RE for tract neighbors + IID RE for tract
    scale.model = T,
    graph = "R:/Project/seattle_rental_market/report/spatial_epi/seatract.graph") +
  f(idqtr, model = "ar1") +
  f(idqtr1, model = "iid") +
  f(idtractqtr, model = "iid")

@@

This website provides information for the paper _Estimating Neighborhood Asking Rents using Scraped Craigslist Rental Listings_. The goal of this research is to a.) understand the temporal dynamics of rent estimates in Seattle and b.) forecast the current quarter's rent levels based off of the prior periods. Throughout the following analyses, the training data are 2017 Q2 up to the prior quarter (i.e. `r as.yearqtr(Sys.Date())-.25`). The test period is a forecast for the current period and includes comparison to the appropriate median rent estimates for data observed so far in this period.

Most graphics include some level of interactivity, usually either hover-over tooltip information or a slider to control various views of the graphic. You can download the source code and data for this project from Github [here](https://github.com/hesscl/smooth-sea)

@@

<br>
<hr>

**Table of Contents**

*  Distribution: density graphic to investigate the distribution of rents among Seattle neighborhoods for each quarter
*  Panel Time-Series: line graphic to show the observed or modeled temporal structure
*  Spatial Time-Series: series of maps to show observed change across time
*  Model fit: tables of model root mean square error (RMSE), mean absolute error (MAE), and deviance information criterion (DIC) across training and test data

@@

```{r index.maps, echo=F}
obs <- sea_f %>%
  mutate(rent = actualRent) %>%
  filter(listingQtr == as.yearqtr(Sys.Date())) %>%
  mutate(var = paste0(as.yearqtr(Sys.Date()), "-to-Date")) %>%
  dplyr::select(GISJOIN, long, lat, group, rent, var, actualRent, spt_Med, NAMELSAD10, n1B)

pred <- sea_f %>%
  mutate(rent = spt_Med) %>%
  filter(listingQtr == as.yearqtr(Sys.Date())) %>%
  mutate(var = paste0(as.yearqtr(Sys.Date()), " Forecast")) %>%
  dplyr::select(GISJOIN, long, lat, group, rent, var, actualRent, spt_Med, NAMELSAD10, n1B)

comb <- bind_rows(obs, pred) %>%
  mutate(tooltip = paste0(NAMELSAD10, "<br>", "Forecast: $", round(spt_Med), "<br>", "Qtr-to-Date: $", round(actualRent), "<br>", "N Qtr-to-Date: ", n1B))
comb <- SharedData$new(comb, ~GISJOIN)

@@

ggplotly(avp_map, tooltip = "label") %>%
  layout(autosize = T, margin = list(b = 50, r = 50, l = 50, t = 50)) %>%
  highlight(on = "plotly_click")
```

@@

```{r, echo=F}
gg_hist <- ggplot(sea_df, aes(x = med1B, group = listingQtr, color = listingQtr)) +
  geom_density() +
  theme_minimal() +
  scale_x_continuous(labels = scales::dollar) +
  scale_color_viridis_d(end = .8) +
  xlab("\nMedian 1B Rent") +
  labs(color = "Quarter")

ggplotly(gg_hist, tooltip = c("color", "x")) %>%
  layout(autosize = T)
```

@@

int_df <- sea_df %>%
  group_by(listingQtr) %>%
  mutate(value = int_Med,
         qtrDev = value-mean(value, na.rm=T),
         model = "B. Quarter FE") %>%
  dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)

ns_df <- sea_df %>%
  group_by(listingQtr) %>%
  mutate(value = ns_Med,
         qtrDev = value-mean(value, na.rm=T),
         model = "C. Non-Spatial RE") %>%
  dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)

nsar1_df <- sea_df %>%
  group_by(listingQtr) %>%
  mutate(value = nsar1_Med,
         qtrDev = value-mean(value, na.rm=T),
         model = "D. Non-Spatial AR(1) RE") %>%
  dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)

bym_df <- sea_df %>%
  group_by(listingQtr) %>%
  mutate(value = bym_Med,
         qtrDev = value-mean(value, na.rm=T),
         model = "E. Spatial RE") %>%
  dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)

spt_df <- sea_df %>%
  group_by(listingQtr) %>%
  mutate(value =  spt_Med,
         qtrDev = value-mean(value, na.rm=T),
         model = "F. Spatiotemporal AR(1)") %>%
  dplyr::select(listingQtr, value, GISJOIN, qtrDev, model)

@@

obs_lineplot <- ggplot(panel_df, aes(x = listingQtr, y = value, group = GISJOIN, label = GISJOIN, color = qtrDev, frame = model)) +
  geom_line(lwd = .25, color = "black", alpha = .5) +
  geom_point() +
  scale_color_gradient2_tableau() +
  scale_y_continuous(labels = scales::dollar) +
  xlab("") +
  ylab("Median 1B Rent\n") +
  theme_minimal() +
  labs(color = "Quarterly\ndeviation")

ggplotly(obs_lineplot, tooltip = c("label", "y")) %>%
  layout(autosize = T) %>%
  animation_opts(frame = 1000, easing = "linear", redraw = FALSE) %>%
  animation_button(hide = T) %>%
  highlight()
```

@@

Column {.tabset}
-------------------------------------

### Observed

```{r spat.series.actual, echo=F}
map_panel_obs <- ggplot(sea_f, aes(x = long, y = lat, group = group, fill = actualRent, frame = listingQtr)) +
  geom_polygon(color = "grey85", lwd = .05) +
  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey80") +
  coord_quickmap() +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_blank()) +
  labs(fill = "Median<br>1B Rent")

ggplotly(map_panel_obs) %>%
  layout(autosize = T) %>%
  animation_opts(transition = 0, redraw = FALSE)
```

### Spatiotemporal AR(1) RE

```{r spat.series.nsar, echo=F}
map_panel_spt <- ggplot(sea_f, aes(x = long, y = lat, group = group, fill = spt_Med, frame = listingQtr)) +
  geom_polygon(color = "grey85", lwd = .05) +
  scale_fill_viridis_c(labels = scales::dollar, na.value = "grey80") +
  coord_quickmap() +
  theme(axis.title = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_blank()) +
  labs(fill = "Median<br>1B Rent")

ggplotly(map_panel_spt) %>%
  layout(autosize = T) %>%
  animation_opts(transition = 0, redraw = FALSE)
```

@@

### Model legend

* int - Quarter fixed intercept
* ns - Non-spatial tract random effect for each tract, quarter fixed intercept
* nsar1 - Non-spatial tract random effect model, AR(1) random effect for quarter, i.i.d. random effect for quarter
* bym - Spatial intrinsic conditional autoregressive (ICAR) tract random-effect with quarter fixed intercept
* spt - Spatial intrinsic conditional autoregressive (ICAR) tract random-effect, AR(1) random effect for quarter, i.i.d. random effect for quarter and tract-quarter

@@

### DIC
```{r, echo=F}
knitr::kable(data.frame("int_DIC" = m.int$dic$dic,
                        "ns_DIC" = m.ns$dic$dic,
                        "nsar1_DIC" = m.nsar1$dic$dic,
                        "bym_DIC" = m.bym$dic$dic,
                        "spt_DIC" = m.spt$dic$dic))
```

Column {.tabset}
-------------------------------------

### Non-Spatial AR(1)

```{r, echo=F}
avp_scatter <- ggplot(sea_df, aes(x = actualRent, y = nsar1_Med, size = n1B, frame = listingQtr)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1) +
  scale_x_continuous(labels = scales::dollar) +
  scale_y_continuous(labels = scales::dollar) +
  xlab("\nActual") +
  ylab("Predicted\n") +
  scale_size_area() +
  coord_equal() +
  theme_minimal()

ggplotly(avp_scatter) %>%
  layout(autosize = T, margin = list(b = 50, r = 50, l = 50, t = 50)) %>%
  animation_opts(1000, easing = "linear", redraw = FALSE)
```

### Spatial RE

```{r, echo=F}
avp_scatter <- ggplot(sea_df, aes(x = actualRent, y = bym_Med, size = n1B, frame = listingQtr)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1) +
  scale_x_continuous(labels = scales::dollar) +
  scale_y_continuous(labels = scales::dollar) +
  xlab("\nActual") +
  ylab("Predicted\n") +
  scale_size_area() +
  coord_equal() +
  theme_minimal()

ggplotly(avp_scatter) %>%
  layout(autosize = T, margin = list(b = 50, r = 50, l = 50, t = 50)) %>%
  animation_opts(1000, easing = "linear", redraw = FALSE)
```

### Spatiotemporal AR(1)

```{r, echo=F}
avp_scatter <- ggplot(sea_df, aes(x = actualRent, y = spt_Med, size = n1B, frame = listingQtr)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1) +
  scale_x_continuous(labels = scales::dollar) +
  scale_y_continuous(labels = scales::dollar) +
  xlab("\nActual") +
  ylab("Predicted\n") +
  scale_size_area() +
  coord_equal() +
  theme_minimal()

ggplotly(avp_scatter) %>%
  layout(autosize = T, margin = list(b = 50, r = 50, l = 50, t = 50)) %>%
  animation_opts(1000, easing = "linear", redraw = FALSE)
```
